---
title: "HW7"
author: "Kayla Dyer"
date: "2025-10-24"
output: word_document
---
## 1. Read in the data set called HW7data.csv. Remember to set your working directory.
```{r}
setwd("~/Desktop/STA 503/Week 7")
data <- read.csv(file = "HW7data.csv")
str(data)
```
## 2. Use the filter( ) function within the tidyverse package to find the number flights that took place on January 1st of that year. There are month (1 - 12) and day (1 - 31) variables. Do NOT print out all the values!
```{r}
library(tidyverse)
jan1_flights <- filter(data, month==1, day==1)
nrow(jan1_flights)
```
## 3. Use the filter( ) function within the tidyverse package to find the number flights that flew to Houston (IAH or HOU values within the dest variable). Do NOT print out all the values!
```{r}
houston_flights <- filter(data, dest=="IAH"|dest=="HOU")
nrow(houston_flights)
```
## 4. Use the filter( ) function within the tidyverse package to find the number of flights weren’t delayed (on arrival or departure) by more than two hours. Here, arr_delay and dep_delay are the two variables of interest. Do NOT print out all the values!
```{r}
on_time_flights <- filter(data, arr_delay<= 120, dep_delay<= 120)
nrow(on_time_flights)
```
## 5. Use the select( ) function within the tidyverse package to write code that prints the first ten observations and only the variables between year and day (inclusive).
```{r}
head(select(data, year:day),10)
```
## 6. Use the mutate( ) function within the tidyverse package to write code that creates a new variable called speed which is equal to (distance/air time)*60. Find the mean and standard deviation of this new variable.
```{r}
data <- mutate(data, speed =distance/ air_time*60)
mean(data$speed, na.rm=TRUE)
sd(data$speed, na.rm=TRUE)
```