---
title: "HW3"
author: "Kayla Dyer"
date: "2025-09-14"
output: word_document
---
## Given vectors in R
```{r}
x <- 2:6
y <- seq(from = 1, to = 13, by = 3)
z <- c(1,2,4,7,11)
names.1 <- c("Kelly", "Jim", "Jeff", "Jennifer", "Daniel", "Drew")
names.2 <- c("Nate", "Christine", "Bob", "Danielle", "Brian", "Monica")
```
## (a) Add 2 to every element in z
```{r}
z + 2
```
## (b) Multiply vectors x and y element-by-element
```{r}
x*y
```
## (c) Print the first, second, and fifth elements of names.1
```{r}
names.1[c(1, 2, 5)]
```
## (d) Print a logical vector that is TRUE if the element of y is equal to 4 and FALSE otherwise
```{r}
y == 4
```
## (e) Combine vectors x and y into one vector
```{r}
c(x, y)
```
## (f) Repeat the second and third elements of x three times
```{r}
rep(x[2:3], times = 3)
```
## (g) Print a logical vector that is TRUE if the element z is less than 7 and FALSE otherwise
```{r}
z < 7
```
## (h) Print the values of y such that those elements were TRUE from the previous logical vector
```{r}
y[z < 7]
```
## (i) Print a logical vector that is TRUE if the element in names.1 comes before the same element in names.2 alphabetically
```{r}
names.1 < names.2
```
## (j) Raise each element of z to the third power
```{r}
z^3
```

