---
title: "HW8"
author: "Kayla Dyer"
date: "2025-10-30"
output: word_document
---
## 1. Read in the data set called HW7data.csv. This is the same data set as the last homework! Load the tidyverse package. Hide the message in the output document using code chunk options.
```{r setup, message=FALSE}
setwd("~/Desktop/STA 503/Week 7")
library(tidyverse)
flights <- read.csv(file = "HW7data.csv")
glimpse(flights)
```
## 2. Use the group by( ) and summarize( ) functions to print the number of flights that took off from each of the three origin airports that year. Print this table in the output document using the kable( ) function from the knitr package.
```{r}
library(knitr)
flights %>%
  group_by(origin) %>%
  summarize(n_flights = n()) %>%
  kable(caption = "Number of flights by origin")
```
## 3. Use all of the functions learned in the last two weeks, along with the pipe command, to find the average speed (see HW 7) of all the planes (tailnum) that took off in April (month) for the three origin airports. Arrange them from fastest to slowest. Print the top ten. Hide the warning message in the output document if you get one. In other words, you are grouping by tailnum and origin and finding the average speed. But only considering all flights that took off in April. And then printing the top 10 fastest.
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)

flights %>%
  filter(month == 4, is.na(air_time) == FALSE, air_time > 0, is.na(distance) == FALSE) %>%
  mutate(speed = (distance / air_time) * 60) %>%                   
  group_by(tailnum, origin) %>%
  summarize(avg_speed = mean(speed, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(avg_speed)) %>%
  slice_head(n = 10) %>%
  kable(caption = "Top 10 fastest tail numbers in April (avg speed, mph)")
```

