---
title: "HW6"
author: "Kayla Dyer"
date: "2025-10-07"
output: word_document
---
## 1. Read in the data set called HW6data.csv. Do not print the data set. Print the structure of the data set. How many observations do we have? How many variables?
```{r}
setwd("~/Desktop/STA 503/Week 6")
data <- read.csv(file = "HW6data.csv")
str(data)
```
## The dataset has 20 observations and 4 variables.

## 2. Print a summary of the variables. Do we have any missing values? If so, how many in each variable?
```{r}
summary(data)
colSums(is.na(data))
```
## Yes there are missing values. There are 3 missing values in variable 2 and 3 missing values in variable 4. 

## 3. Print the mean and standard deviation of variables var.1 and var.2.
## var.1
```{r}
mean(data$var.1, na.rm = TRUE)
sd(data$var.1, na.rm = TRUE)
```
##var.2
```{r}
mean(data$var.2, na.rm = TRUE)
sd(data$var.2, na.rm = TRUE)
```
## 4. Print the distribution (the table) of relative frequencies for variable var.3
```{r}
table(data$var.3)
prop.table(table(data$var.3))
```
## 5.Print the two-way table of frequencies with var.3 as the rows and var.4 as the columns. Explain why all the numbers don’t add up to 20 (the number of observations).
```{r}
table(data$var.3, data$var.4)
```
## The numbers don't add up to 20 because 3 observations in the dataset have missing values. 
## 6. Make a scatterplot with var.1 as your explanatory variable and var.2 as your response variable. Add the OLS regression line to this scatterplot.
```{r}
plot(data$var.1, data$var.2)
lm.fit <- lm(var.2 ~ var.1, data = data)
abline(lm.fit)
```

## 7. Using the kable function and inline code, print the coefficient table from the lm( ) function. In other words, your output document should have a table outside of the code chunk and no R code should be included.
```{r,echo=FALSE}
lm.fit <- lm(var.2 ~ var.1, data = data)
coef_table <- summary(lm.fit)$coefficients
knitr::kable(coef_table, digits = 4,
             caption = "Coefficient Table for Linear Model: var.2 ~ var.1")
```
## 8. Make side-by-side horizontal boxplots of var.1 across the different levels of var.3
```{r}
boxplot(var.1 ~ var.3, data = data,
        horizontal = TRUE,
        main = "Boxplots of var.1 by var.3",
        xlab = "var.1",
        ylab = "var.3")
```
## 9. Recreate the plot on the following page in R. Here, the A, B, and C come from var.4. The Yes,and No come from variable var.3
```{r}
counts.2 <- table(data$var.3, data$var.4)
barplot(height = counts.2,
        beside = TRUE,
        legend = rownames(counts.2),
        main = "Bar Plot of var.3 by var.4",
        xlab = "var.3",
        ylab = "Frequency")
```

